### 2013-07-04 proposal.tex

\documentclass[10pt,a4paper]{article}

\usepackage{fullpage}

\title{Designing a Unified BBC Search\\ \vspace{2 mm}{\large Systems Design Coursework Proposal}}
\author{Ross Fenning}

\begin{document}

\maketitle

\input{proposal-contents}

\end{document}