### 2013-08-24 coursework.tex
Added bibliography and sort of finished the first chapter.

\documentclass[a4paper]{report}

\title{Designing a Search for the BBC}
\author{Ross Fenning}

\begin{document}

\maketitle

\tableofcontents

\chapter{The Problem}

\section{Rationale}

The BBC has been publishing content on the World Wide Web since the
mid 1990s and since then the amount and the diversity has increased
exponentially. Large websites or indeed the web as a whole would not have
been usable nor useful without the rise in quality of web search engines to
help people find content based on keywords or phases that describe or
appear in the sought after pages.

A large challenge for any web search application is to provide a common
interface and set of user interactions that can equally index, search
and link to a diverse range of types of information -- be it in the form
of text, images, video or games. A more recent challenge has been to
achieve this in a near-real time way to catch up with the rapid rate at
which content is added to the web (particulary from microblogging websites
such as Twitter).

Whilst there are successful general web search engines such a Google that
will find any piece of content, there is a perceived need for a
BBC-specific search that makes better use of the internal knowledge the BBC
has about its own content.

\section{Objectives}

The purpose of this report is to start the high-level design -- and perhaps
look at the lower levels of one or two aspects -- of a search application
specific to BBC content. The intent is to provide a consistent user interface
that allows the audience to find equally news articles, sport results,
TV catch-up, education resources and everything else the BBC produces.

The target audience for the BBC is effectively the entire population of the UK
and amongst those that do make use of BBC services, there is much diversity
of needs, preferences and technical ability. It is clear it is no small
task to design a search-based discovery mechanism of millions of diverse
pieces of content aimed at millions of diverse people.

A suitable approach for tackling real-word problems is \emph{Soft Systems Methodology} (SSM)
\cite{checkland2006learning}, which will allows us to stand back from
an ontological approach of defining what the search system \emph{is} or
\emph{comprises} and instead take an \emph{epistemological} view of search
as a system. With this view, we could consider a system that holistically
transforms members of the public's desires to find online content into
the consumption of that content -- whether those desires are \emph{precise}
(e.g. they want an exact article known by headline they saw earlier or a
particular programme they missed on television) or those desires are
\emph{fuzzy} (e.g. news about a certain topic, any comedy programme, learning
materials about the Industrial Revolution).

Checkland \cite{checkland1990soft} decribed a \emph{Rich Picture} approach to
representing a problem situation early in SSM approaches.
Given the size and complexity of the
search system as a whole, a useful initial step is to create such an informal
representation of what is known about the problem. Figure~\ref{rich-picture}
shows what I know of the audience, search and most BBC online content areas.
Note that not all areas are covered and a strong emphasis is placed on TV
catch-up (e.g. via the iPlayer product). Radio catch-up is not mentioned
as it shares a lot of similarity with television in terms of use and any
differences are out of scope for this design.

\chapter{Design}

\section{Use cases}

\chapter{Analysis}

\chapter{Evaluation}

\appendix
\chapter{Proposal}

\input{proposal-contents}

\bibliographystyle{cell}
\bibliography{bibtex}

\end{document}